---
title: "Simulate with modified parameter values"
output:
rmarkdown::html_vignette:
    toc: true
Suggests: markdown
VignetteBuilder: knitr
vignette: >
  %\VignetteIndexEntry{VaryParams}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
- \usepackage{ae}
---

```{r,include = FALSE}
library(data.table)
library(NMdata)
library(ggplot2)
library(NMsim)
##knitr::opts_chunk$set(dev = "cairo_pdf")
knitr::opts_chunk$set(
                      collapse = TRUE
                     ,comment = "#>"
                     ,fig.width=7
                     ,cache=FALSE
                  )

NMdataConf(as.fun="data.table",
           dir.sims="~/NMsim_vignette",
           dir.res="simulate-results",
           allow.unknown=TRUE ## necessary for dir.sims and dir.res
           ## until NMdata 0.1.5
           )

theme_set(theme_bw()+theme(legend.position = "bottom"))

## this changes data.table syntax. I think we can do without.
## knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
run.simuls <- FALSE
```
Built `r Sys.Date()` using NMsim `r packageVersion("NMsim")`.

## Objectives
This vignettes aims at enabling you to use `NMsim` for the following purposes

* Simulation with parameters modified from the estimated values

## Vary parameter values
Sometimes we want to simulate with some modification to the estimated model. NMsim can make such user-specified modifications to the model before simulating through the `list.sections` argument. 

The SAD study was run with a fast solution formulation. We want to see how a slower absorption rate would affect the PK prediction for the multiple dose regimen. In the model estimate, `TVKA=2.17`. We now try with a four times slower absorption:

Notice, until NMsim 0.1.0, the `modify.model` argument was called `modify.model`. The helper functions `add` and `overwrite` were introduced with NMsim 0.1.3. Until then, custom modification functions had to be provided like described in the section "Edit control stream using customized functions".

```{r,include=FALSE,eval=TRUE}
file.mod <- system.file("examples/nonmem/xgxr021.mod",package="NMsim")
dat.sim <- read.csv(system.file("examples/derived/dat_sim1.csv",package="NMsim"))
```

```{r,eval=FALSE}
file.mod <- system.file("examples/nonmem/xgxr021.mod",package="NMsim")
dat.sim <- read.csv(system.file("examples/derived/dat_sim1.csv",package="NMsim"))

simres <- NMsim(file.mod=file.mod
               ,data=dat.sim
               ,dir.sims="~/NMsim_vignette" ## where to store simulation files
               ,name.sim="original"
               ,seed.R=12345
                )

simres.slowabs <- NMsim(file.mod=file.mod,
                        data=dat.sim
                       ,dir.sims="~/NMsim_vignette" ## where to store simulation files
                       ,name.sim="slower_abs"
                       ,seed.R=12345
                       ##,modfiy.model=list(PK=function(x)c(x,"TVKA=TVKA/4","KA=KA/4"))
                       ,modify.model=list(PK=add("TVKA=TVKA/4","KA=KA/4"))
                        )

```
```{r,include=FALSE,eval=TRUE}
simres <- NMreadSim("simulate-results/xgxr021_original_MetaData.rds")
simres.slowabs <- NMreadSim("simulate-results/xgxr021_slower_abs_MetaData.rds")


## file.fst <- "simulate-results/simres_slowabs.fst"
## if(run.simuls){
##     write_fst(simres.slowabs,path=file.fst)
## } else {
##     simres.slowabs <- read_fst(file.fst,as.data.table=TRUE)
## }
## simres <- read_fst("simulate-results/simres.fst",as.data.table=TRUE)
```

```{r,eval=TRUE}
rbind(simres,simres.slowabs) |>
    ggplot(aes(TIME,PRED,colour=model))+
    geom_line()

```

We used `modify.model` to modify the `$PK` section. We used it to
append two lines. We could use it to modify any section in the model,
and we could essentially do any modification. However, appending to
`$PK` or `$PRED` is simple and often both robust and flexible enough.

That was a very spcific analysis of one specific `KA` value. It is
often more convenient to control the numeric changes to the model
using the simulation input data set rather than hard-coding numerical
values into `modify.model`. The following tries a number of fold
changes to `KA`.

```{r,expand-input-dt,eval=TRUE}
setDT(dat.sim)
## this repeats dat.sim for each of the values in KASCAL - effectively
## an outer join
dat.sim.varka <- dat.sim[,data.table(KASCALE=c(1,4,10)),by=dat.sim]
dat.sim.varka[,ID:=.GRP,by=.(KASCALE,ID)]
setorder(dat.sim.varka,ID,TIME,EVID)
```

```{r,include=TRUE,eval=FALSE}
simres.varka <- NMsim(file.mod=file.mod,
                      data=dat.sim.varka
                     ,dir.sims="~/NMsim_vignette" ## where to store simulation files
                     ,name.sim="varka"
                     ,seed=12345
                     ## ,modify.model=list(PK=function(x)c(x,"TVKA=TVKA/KASCALE","KA=KA/KASCALE"))
                     ,modify.model=list(PK=add(c("TVKA=TVKA/KASCALE","KA=KA/KASCALE")))
                      )

```
```{r,include=FALSE,eval=TRUE}
## file.fst <- "simulate-results/simres_varka.fst"
## if(run.simuls){
##     write_fst(simres.varka,path=file.fst)
## } else {
##     simres.varka <- read_fst(file.fst,as.data.table=TRUE)
## }
simres.varka <- NMreadSim("simulate-results/xgxr021_varka_MetaData.rds")
```

```{r,eval=TRUE}
ggplot(simres.varka[simres.varka$EVID==2,],aes(TIME,PRED,colour=factor(KASCALE)))+
geom_line()+
labs(colour="Fold absorption prolongation")
```

## Edit control stream using customized functions
The `modify.model` argument can be used to apply customized functions
to control stream sections. In the example above, we used 
```{r,eval=FALSE}
modify.model=list(PK=add(c("TVKA=TVKA/KASCALE","KA=KA/KASCALE")))
```
This is exactly the same as creating a custom function that adds those two lines (text strings to a vector of text strings each representing a line in the control stream)

```{r,eval=FALSE}
modify.model=list(PK=function(x)c(x,"TVKA=TVKA/KASCALE","KA=KA/KASCALE"))
```

We can use a difference custom function to modify `TVKA` when it is first defined in
the control stream. The following uses a regular expression to substitute 




<!-- ## Using NMsim to validate reimplementations of models -->

```{r,include=TRUE,eval=FALSE}
simres.varka2 <- NMsim(file.mod=file.mod,
                      data=dat.sim.varka
                     ,dir.sims="~/NMsim_vignette" ## where to store simulation files
                     ,name.sim="varka2"
                     ,seed.R=12345
                     ,modify.model=list(PK=function(x) {
                         ## identify line number if first definition of TVKA
                         idx.line1.TVKA <- min(grep(" *TVKA *=",x))
                         ## add /KASCALE after dropping potential comments
                         x[idx.line1.TVKA] <- paste(sub(";.*","",x[idx.line1.TVKA]),"/KASCALE")
                         x})
                      )

```

```{r,include=FALSE,eval=TRUE}
## file.fst <- "simulate-results/simres_varka.fst"
## if(run.simuls){
##     write_fst(simres.varka,path=file.fst)
## } else {
##     simres.varka <- read_fst(file.fst,as.data.table=TRUE)
## }
simres.varka2 <- NMreadSim("simulate-results/xgxr021_varka2_MetaData.rds")
```

```{r}
simres.both <- rbind(simres.varka[,method:="add()"],
                     simres.varka2[,method:="custom/sub()"]
                     )

ggplot(simres.both[EVID==2],aes(TIME,PRED,colour=factor(KASCALE)))+
geom_line()+
    labs(colour="Fold absorption prolongation")+
    facet_wrap(~method)
```

