---
title: "Creation of Simulation Data Sets"
output:
rmarkdown::html_vignette:
    toc: true
    code_folding: show
Suggests: markdown
VignetteBuilder: knitr
vignette: >
  %\VignetteIndexEntry{06dsCreate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
- \usepackage{ae}
---

```{r,include = FALSE}
##knitr::opts_chunk$set(dev = "cairo_pdf")
knitr::opts_chunk$set(
                      collapse = TRUE
                     ,comment = "#>"
                     ,fig.width=7
                     ,cache=FALSE
                  )

## this changes data.table syntax. I think we can do without.
## knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

```{r,setup,include=F}
## library(devtools)
## unloadNamespace("NMsim")
## unloadNamespace("NMdata")

## load_all("~/wdirs/NMdata")
## load_all()
library(NMsim)
library(data.table)
library(NMdata)
library(ggplot2)
library(patchwork)
library(tracee)
library(fst)
library(knitr)
## NMdataConf(path.nonmem="/opt/NONMEM/nm75/run/nmfe75")
## NMdataConf(path.nonmem="/opt/nonmem/nm751/run/nmfe75")
## NMdataConf(dir.psn=NULL)

theme_set(theme_bw())
this.script <- "NMsim-DataCreate.Rmd"
writeOutput <- TRUE
file.project <- function(...)file.path(system.file("examples",package="NMsim"),...)
## file.project <- function(...)file.path("../inst/examples",...)
## file.project <- function(...)file.path("~/wdirs/NMsim/inst/examples",...)


run.simuls <- FALSE
NMdataConf(as.fun="data.table")
```

Built `r Sys.Date()` using NMsim `r packageVersion("NMsim")`.



### A basic simulation data set

As long as a simulation data set is in a `data.frame` structure with columns and values compatible with the Nonmem model(s) to be simulated, it does not matter to `NMsim` how it was created. However, `NMsim` and `NMdata` do provide powerful methods to easily create such data sets.

As an example we
create a regimen with a loading dose of 300 mg followed by 150 QD for
6 days. We dose into compartment 1, and we want to simulate samples in the second compartment. These numbers depend on the model which the data set is intended to be used with.

`NMcreateDoses()` is a flexible function that creates dosing records based on a concise syntax. We add a label to the regimen right away.

```{r,dsCreateSim,include=TRUE}
### multiple dose regimens with loading are easily created with NMcreateDoses too
## We use ADDL+II (either method easy)
doses <- NMcreateDoses(TIME=c(0,24),AMT=c(300,150),addl=data.frame(ADDL=c(0,5),II=c(0,24)),CMT=1)
doses <- transform(doses,trt="300 mg then 150 mg QD")
## Notice, the ID and MDV columns are included
doses
```

Now we add the sample records using `addEVID2()`. 

```{r}
## Add simulation records - longer for QD regimens
dat.sim <- addEVID2(doses,time.sim=0:(24*7),CMT=2)
```

`dat.sim` is now a valid simulation data set with one subject. However, even though `addEVID2()` does try to order the data in a meaningful way, it is recommended to always manually order the data set. We use `data.table`'s `setorder()`. `dplyr::arrange` can just as well be used. A row identifier (counter) can make post-processing easier, so we add that too.

```{r}
## sort data set 
setorder(dat.sim,ID,TIME,EVID)
## Adding a row identifier (generally not necessary but recommended)
dat.sim$ROW <- 1:nrow(dat.sim)
```


```{r,save-datsim,eval=FALSE,include=FALSE}
## NMwriteData(dat.sim,file="simulate-results/dat_sim.csv",genText=FALSE)
NMwriteData(dat.sim,file="../inst/examples/derived/dat_sim1.csv",genText=FALSE)
```

`NMsim` does not include any plotting functionality, but here is a simple way to show dosing amounts and sample times. `NMdata::NMexpandDoses()` is used to _expand_ the doses coded with `ADDL`/`II` in order to get a data row to plot for each dose. We also take the sum of the amounts by time point in case doses are simultaneous. 

```{r,fig.height=2}
dtplot <- NMdata::NMexpandDoses(dat.sim,as.fun="data.table")
dtplot <- dtplot[,.(AMT=sum(AMT)),by=.(ID,CMT,TIME,EVID)]

ggplot(dtplot,aes(TIME,factor(CMT),colour=factor(EVID)))+
    geom_point(data=function(x)x[EVID==1],aes(size=AMT))+
    geom_point(data=function(x)x[EVID==2],shape="|")+
    labs(x="Time (hours)",y="Compartment")+
    theme(legend.position="bottom")

````


A brief overview of the number of events broken down by event type `EVID` and dose amount `AMT`:

```{r,echo=FALSE}
NMexpandDoses(dat.sim)[,.(Nrows=.N),keyby=.(trt,CMT,EVID,AMT)] |>
    kable()
```

Showing the top five rows for understanding what the data now looks like. Notice that the following are _not_ issues:

- Data contains a mix of numeric and non-numeric columns
- Columns are not sorted in Nonmem-friendly style with non-numeric columns to the right


```{r,echo=FALSE}
dat.sim[1:5,]
```


Finally, We check the simulation data set for various potential issues in
Nonmem data sets using `NMdata::NMcheckData` and summarize the number
of doses and observations:

```{r}
NMdata::NMcheckData(dat.sim,type.data="sim")
```

### Add time after previous dose and related information

Just like for preparation of data set with observed data, time after previous dose may be the relevant time to analyze the simulation results against. We use `NMdata::addTAPD()` to add this automatically.

```{r,addTAPD}
dat.sim2 <- addTAPD(dat.sim)
head(dat.sim2)
```

Here is something important to decide for simulations. Notice `TAPD` for the sample at `TIME==0`.
`addTAPD` does not use the order of the data set to determine the time-order or the records. The default behavior of `addTAPD` is to treat a sample taken at the exact same time as a dose as a pre-dose. If instead we want them to be considered post-dose, we have to specify how to order `EVID` numbers.

```{r,addTAPD2}
## order.evid=c(1,2) means doses are ordered before EVID=2 records
dat.sim2 <- addTAPD(dat.sim,order.evid=c(1,2))
## now the TIME=0 sample has TAPD=0
head(dat.sim2)
```

`addTAPD` uses `NMdata::NMexpandDoses` to make sure all dosing times are considered. See `?NMdata::addTAPD` for what the other created columns mean and for many useful features.

### Multiple endpoints (e.g. parent and metabolite)

Pass a data.frame to `addEVID2`'s CMT argument to include multiple endpoints. 

```{r}
addEVID2(doses,CMT=data.frame(CMT=c(2,3),DVID=c("Parent","Metabolite")),time.sim=1:2)
```


### Cohort-dependent or individual sampling schemes
Same way as for the `CMT` argument, `time.sim` can also be a `data.frame`. If it contains a covariate found in the doses data, the added simulation times will be merged on accordingly. You can use say a cohort identifier, or it could be `ID` which allows you to reuse (all or parts of) the observed sample times.

### Other features
As the name implies the default of function is to use `EVID=2` (which means they are neither doses, samples, nor resetting events) for these records. Should you want to change that (maybe to `EVID=0`), use the `EVID` argument.
