---
title: "VPC simulations"
output:
rmarkdown::html_vignette:
    toc: true
Suggests: markdown
VignetteBuilder: knitr
vignette: >
  %\VignetteIndexEntry{02VPC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
- \usepackage{ae}
---

```{r,include = FALSE}
## library(tidyvpc)
library(NMdata)
library(fst)
library(NMsim)
##knitr::opts_chunk$set(dev = "cairo_pdf")
knitr::opts_chunk$set(
                      collapse = TRUE
                     ,comment = "#>"
                     ,fig.width=7
                     ,cache=FALSE
                  )

## NMdataConf(dir.psn="/opt/psn")

## this changes data.table syntax. I think we can do without.
## knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
run.simuls <- FALSE
```

## Simulations for Visual Predictive Checks (VPC)
This vignette shows how to generate simulations for generation of VPC
plots. While `NMsim` does not include any functionality for summarizing quantiles or plotting, it provides powerful ways to obtain the simulated data needed. We shall see how the `tidyvpc` package easily creates VPC plots based on the simulation results.

## Default option: reuse estimation data for simulation
Normally, the two main arguments to NMsim are the path to the input
control stream (`file.mod`) and the simulation input data set
(`data`). But if we leave out the the `data` argument, NMsim will
re-use the estimation data for the simulation. That is the simulation
we need for a VPC. We will use an example model included with NMsim:

```{r,eval=TRUE}


file.project <- function(...)file.path(system.file("examples",package="NMsim"),...)
file.mod <- file.project("nonmem/xgxr032.mod")

```

```{r,eval=FALSE}
set.seed(43)
## notice the data argument is not used.
simres.vpc <- NMsim(file.mod,
                    table.vars=c("PRED","IPRED", "Y"),
                    dir.sims="testOutput",
                    name.sim="vpc_01"
                   ,nsims=500
                    )

```

The performed simulation is similar to the one produced by the `VPC` function in `PSN`. However, there are some important differences.

* The simulation results are automatically read into R.

* No postprocessing of the results is being done by `NMsim` 

* The `table.vars` argument allows the user to narrow down the variables to be written to disk. This can speed up the simulation considerably and reduce the amount of disk space the Nonmem simulation results require.




## Plotting using `tidyvpc`
As mentioned, `NMsim` does not postprocess the simulation for generation of a VPC plot, nor does it offter any plotting functions. The R package called `tidyvpc` offer those two things and is moreover implemented in `data.table`, so it's fast. The following simple code shows how to get from the results from `NMsim` to the VPC plot with `tidyvpc`.


```{r,include=FALSE,eval=TRUE}
simres.vpc <- NMreadSim(file.res)
```


```{r,eval=TRUE}
library(ggplot2)
library(tidyvpc)
library(NMdata)
## read the data as it was used in the Nonmem model
res <- NMscanData(file.mod,as.fun="data.table")
## only plot observation events from estimation data set
data.obs <- res[EVID==0]
## Only plot simulated observation events
data.sim <- subset(simres.vpc,EVID==0)

## run vpc
vpc1 <-
    observed(data.obs, x = TIME, y = DV) |>
    simulated(data.sim, y = Y) |>
    stratify(~DOSE) |>
    binning(bin = "ntile", nbins = 9) |>
    vpcstats()


plot(vpc1)

```


### Use a different input data set
In the first example we used the exact same data as was used for the
estimation. This is a common way to produce a VPC, and we saw the
advantage that the user does not risk making mistakes in preparing the
data set for the simulations. However, it may be of interest to
include additional data or even a different data set in the
simulation. It could be including data points that were excluded in
estimation (like samples below the quantification limit) or a separate
study that was not included in the model. All you have to do is to
read the data you want and provide it in `NMsim`'s `data` argument.

### Make use of the cluster
We will repeat the same as above, but now 500 times (`nsims`). We make
use of a few more arguments for efficiency. `sge` means that the jobs
will be sent to the cluster. The `nc` argument is now used meaning
only one core will be used per job. If each node on the cluster has 16
cores, this could engage 500/16 ~ 32 nodes in parallel, with all jobs
executed at the same time. We supply the path to the Nonmem executable. With PSN this should work without specifying the Nonmem path, but PSN for some reason takes more time submitting the jobs to the cluster. If nodes are available, the following simulation should not take more than a couple of minutes to execute.

```{r,eval=TRUE}
file.res <- "simulate-results/simpaths-vpc.rds"
```

```{r,eval=FALSE}
set.seed(43)
## notice the data argument is not used.
sim.vpc.sge <- NMsim(file.mod,
                     table.vars=c("PRED","IPRED", "Y"),
                     dir.sims="testOutput",
                     name.sim="vpc_01"
                    ,nsims=500
                    ,sge=TRUE
                    ## ,path.nonmem="/opt/nonmem/nm751/run/nmfe75"
                    ,path.nonmem="/opt/NONMEM/nm75/run/nmfe75"
                    ,file.res=file.res
                     )

```

