---
title: "Simulation of typical subjects"
output:
rmarkdown::html_vignette:
    toc: true
    code_folding: show
Suggests: markdown
VignetteBuilder: knitr
vignette: >
  %\VignetteIndexEntry{Typical}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
- \usepackage{ae}
---

```{r,include = FALSE}
##knitr::opts_chunk$set(dev = "cairo_pdf")
knitr::opts_chunk$set(
                      collapse = TRUE
                     ,comment = "#>"
                     ,fig.width=7
                     ,cache=FALSE
                  )

## this changes data.table syntax. I think we can do without.
## knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

```{r,setup,include=F}
## library(devtools)
## unloadNamespace("NMsim")
## unloadNamespace("NMdata")

## load_all("~/wdirs/NMdata")
## load_all()
library(NMsim)
library(data.table)
library(NMdata)
## library(dplyr)
library(tibble)
library(ggplot2)
library(patchwork)
library(tracee)
## library(tidyr)
library(fst)
library(knitr)
## NMdataConf(path.nonmem="/opt/NONMEM/nm75/run/nmfe75")
## NMdataConf(path.nonmem="/opt/nonmem/nm751/run/nmfe75")
## NMdataConf(dir.psn=NULL)

theme_set(theme_bw())
this.script <- "NMsim-simulate.Rmd"
writeOutput <- TRUE
file.project <- function(...)file.path(system.file("examples",package="NMsim"),...)
## file.project <- function(...)file.path("../inst/examples",...)
## file.project <- function(...)file.path("~/wdirs/NMsim/inst/examples",...)

found.files <- list.files(file.project("nonmem/NMsim"),pattern="noname\\.(lst|xml|ext|cov|cor|coi|phi|msf|msfi|msfo|tab)",full.names=TRUE)
unlink(found.files)

run.simuls <- FALSE
NMdataConf(as.fun="data.table")
```


## Simulation of a typical subject
A typical subject is here understood as a subject without random
effects, i.e. all ETA's equal zero. It is important to realize that
"typical" does not have to do with covariates which the user will
still need to control in the model, in the simulation input data, or
by a combination of these. Getting `NMsim` to run with all ETA's
equaling zero is this easy:

```{r,eval=FALSE}
simres.typ <- NMsim(file.mod=file.mod,
                    data=dat.sim,
                    name.sim="typSubj",
                    method.sim=NMsim_typical)

```

```{r,include=FALSE,eval=TRUE}
file.fst <- "simulate-results/simres_typ.fst"
if(run.simuls){
    write_fst(simres.typ,path=file.fst)
} else {
    simres.typ <- read_fst(file.fst,as.data.table=TRUE)
}
```


In the first simulation we used `PRED` from the default simulation
method to get a typical subject simulation. That will work in many
cases, but that depends on the model. The way to run a simulation with
all ETA's set to 0 is using `method.sim=NMsim_typical`.

```{r,eval=TRUE}
p.typ <- ggplot(simres.typ,aes(TIME,IPRED,color=trt))+geom_line()+
    geom_line(aes(y=PRED))
p.typ

```
