---
title: "NMsim Configuration"
output:
rmarkdown::html_vignette:
    toc: true
Suggests: markdown
VignetteBuilder: knitr
vignette: >
  %\VignetteIndexEntry{Configuration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
- \usepackage{ae}
---

```{r,include = FALSE}
##knitr::opts_chunk$set(dev = "cairo_pdf")
knitr::opts_chunk$set(
                      collapse = TRUE
                     ,comment = "#>"
                     ,fig.width=7
                     ,cache=FALSE
                  )

## this changes data.table syntax. I think we can do without.
## knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

## Objectives
This vignettes aims at enabling you to

* Configure `NMsim` to use `PSN` or methods provided by `NMsim` to update Nonmem control stream initial values and to run Nonmem

* Understand pros and cons of using `PSN` vs. methods provided by `NMsim`.


NMsim relies on Nonmem for running simulations. If your system does
not meet the requirements described in this document, it does not mean
that NMsim can't work. It can still put together your Nonmem
simulations and make everything ready for Nonmem to be executed. Once
you have run Nonmem, you can then use `NMdata::NMscanData()` to
collect the results.

## Requirements
Currently, NMsim only fully supports systems that run Nonmem on
linux. Windows support is experimental.

Notice, the important thing is which system is running Nonmem, not
what system is running R. So if you run R on a windows system and can
run Nonmem through an ssh tunnel, NMsim should still be able to
work. If you run Nonmem on Windows and would like to get NMsim to
work, reach out and we will figure it out.

NMsim functions and their arguments often refer to input control
streams by `.mod`. It should work independently of the file name
extension of the input control streams. The estimate files (`.ext` and
if known subjects simulated `.phi`) are currently expected to carry
the same file name but with those file name extension.

## Configuration of NMsim
Below this section, some background discussion is provided to understand which methods to choose. No matter what you prefer to use, the best is to set up NMsim to be able to use both Nonmem and (if available) PSN. Then you have the flexibility to switch between methods as preferred. 

### Specify the Nonmem paths
For NMsim to run Nonmem, it needs to know where to find the Nonmem
executable.

The easiest way to do configure this is through NMdata's configuration
function. Say you want to run Nonmem with `/opt/NONMEM/nm75/run/nmfe`,
insert this after loading `NMdata` in the beginning of your script

```{r,eval=FALSE}
NMdataConf(path.nonmem="/opt/NONMEM/nm75/run/nmfe75")
```

On Windows, the executable has a `.bat` extension. The path could look be

```{r,eval=FALSE}
NMdataConf(path.nonmem ="C:/nm75g64/run/nmfe75.bat")
```

If you normally use PSN as your Nonmem interface, and you therefore do not know where Nonmem is installed, you can check this using PSN. The following command should give you the Nonmem installation paths that PSN is configured with. However, you likely still need to add the last piece of the path from the installation directory to the nonmem binary.

```
psn -nm_versions
```

### Specify the PSN installation path
If PSN is available, and you can run `execute` and `update_inits` in a
terminal, you don't need to configure anything. If you have PSN
installed, but you have to provide the paths to those two executables
when running them (something like `/opt/PSN/execute run1.mod`), you
will have to tell NMsim where to find them. In this case, the easiest
is loading NMdata and then running:

```{r,eval=FALSE}
NMdataConf(dir.psn="/opt/PSN")
```




## PSN or not?
### PSN or not for Nonmem execution
NMsim can call Nonmem using PSN's `execute` or through similar
functionality included in NMsim. We will refer to those methods by
the way they are referred to in the `NMsim()` function argument that
controls which one is used, i.e. `method.execute="psn"` and
`method.execute="nmsim"`. 

For simulations, `method.execute="psn"` does not provide advantages
over `method.execute="nmsim"`. In contrast, there are simulation types
that will only work with `method.execute="nmsim"`. You need to tell
NMsim where to find Nonmem (by setting `path.nonmem`) for this to work.

Nonmem execution with `NMexec()` (typically, for estimation and not
simulation) is a slightly different discussion. See `?NMexec` for
that.

### PSN or not for updating initial values
NMsim needs to update initial values using the model estimate prior to
running a simulation. PSN provides the `update_inits` function to do
this, and if available, this is the default method in NMsim. Again, we
refer to this by the argument name, `method.update.inits="psn"`.

NMsim also provides a similar functionality internally, referred to as
`method.update.inits="NMsim"`. This will be used if PSN is not found -
or if you ask for it. 

`method.update.inits="psn"` has two advantages over the NMsim-provided
method. It is widely used, and it keeps comments in affected control
stream sections (`$THETA`, `$OMEGA`,
`$SIGMA`). `method.update.inits="nmsim"` will drop all comments in
affected sections. I have in the past seen issues with some methods to
update `$OMEGA` sections with off-diagonal elements
(`BLOCK`). `method.update.inits="nmsim"` takes a very simple
approach. Reading the `OMEGA` matrix from the `.ext` file, this is not
relying on interpreting Nonmem code at all. The support of systems
without `PSN` would be the main reason one would choose
`method.update.inits="nmsim"` over `method.update.inits="psn"`.

In order to make use of `method.update.inits="psn"`, you need to make
sure NMsim can find PSN.
