---
title: "Simulation of New Subjects"
output:
rmarkdown::html_vignette:
    toc: true
    code_folding: show
Suggests: markdown
VignetteBuilder: knitr
vignette: >
  %\VignetteIndexEntry{01BASIC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
- \usepackage{ae}
---

```{r,include = FALSE}
##knitr::opts_chunk$set(dev = "cairo_pdf")
knitr::opts_chunk$set(
                      collapse = TRUE
                     ,comment = "#>"
                     ,fig.width=7
                     ,cache=FALSE
                  )

## this changes data.table syntax. I think we can do without.
## knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

```{r,setup,include=F}
## library(devtools)
## unloadNamespace("NMsim")
## unloadNamespace("NMdata")

## load_all("~/wdirs/NMdata")
## load_all()
library(NMsim)
library(data.table)
library(NMdata)
## library(dplyr)
## library(tibble)
library(ggplot2)
library(patchwork)
library(tracee)
## library(tidyr)
library(fst)
library(knitr)
## NMdataConf(path.nonmem="/opt/NONMEM/nm75/run/nmfe75")
## NMdataConf(path.nonmem="/opt/nonmem/nm751/run/nmfe75")
## NMdataConf(dir.psn=NULL)

theme_set(theme_bw())
this.script <- "NMsim-simulate.Rmd"
writeOutput <- TRUE
file.project <- function(...)file.path(system.file("examples",package="NMsim"),...)
## file.project <- function(...)file.path("../inst/examples",...)
## file.project <- function(...)file.path("~/wdirs/NMsim/inst/examples",...)

found.files <- list.files(file.project("nonmem/NMsim"),pattern="noname\\.(lst|xml|ext|cov|cor|coi|phi|msf|msfi|msfo|tab)",full.names=TRUE)
unlink(found.files)

run.simuls <- FALSE
NMdataConf(as.fun="data.table")

NMdataConf(dir.sims="~/NMsim_vignette",
           dir.res="simulate-results",
           allow.unknown=TRUE ## necessary for dir.sims and dir.res
           ## until NMdata 0.1.5
           )
```


Built `r Sys.Date()` using NMsim `r packageVersion("NMsim")`.


## Objectives
This vignettes aims at enabling you to

* Use `NMsim` to simulate Nonmem models with a given input data set

* Distinguish between and perform the most common types of simulations: 
- new subjects, 

* Simulate multiple new subjects and derive prediction intervals

* Simulate more than one Nonmem model in one `NMsim()` function call

* Important arguments

* Speed up NMsim by avoiding large table statements

* Add residual variability to an already performed model simulation
  using `NMsim`.


## Prerequisites
You should have configured `NMsim` with the path to the Nonmem
installation and maybe also PSN. See
[`NMsim-config.html`](https://philipdelff.github.io/NMsim/articles/NMsim-config.html). Don't
worry - it is very easy.



## Estimation based on single dose, simulation of multiple doses
The situation is like this: We collected PK and PD data from a single ascending dose 
trial on a drug candidate. A PK model was estimated using Nonmem. We have on file
the model input and output control streams (here with extensions `.mod`
and `.lst` respetively), parameter estimates (`.ext`). 
<!-- The following plot shows the data for no other reason than pointing out clearly that the data informing the model was single dose, and we will be simulating a different dosing regimen. -->

```{r,include=FALSE,echo=FALSE}
file.mod <- file.project("nonmem/xgxr021.mod")
res <- NMscanInput(file.mod,quiet=TRUE)
ggplot(res[!is.na(DV)],aes(TIME,DV))+
    geom_point()+
    facet_wrap(~trtact)

```


We want to predict concentrations in a multiple dose regimen. This is a regimen that we have not
studied in clinical trials so far, and we have decided to use
population PK simulations for this purpose. 

### The Simulation data set
You can create a Nonmem-compatible simulation data set however you want. We will keep that as a separate topic and read one that was already created in [`NMsim-DataCreate.html`](https://philipdelff.github.io/NMsim/articles/NMsim-DataCreate.html) using tools provided with NMsim to make that task simpler and faster to do:
```{r}
dat.sim <- read_fst(path="simulate-results/dat_sim.fst")
```

## Simulation of a new subject
This is the first time we are using `NMsim`, and we just want to try
the simplest thing we can think of. Simulate a new subject on the
considerd multiple dose regimen with our estimated PK model from the
single dose study.


```{r,sim-simplest,eval=FALSE}
file.mod <- file.project("nonmem/xgxr021.mod")

simres <- NMsim(file.mod=file.mod,
                data=dat.sim)
```

```{r,include=FALSE,eval=TRUE}
## file.fst <- "simulate-results/simres.fst"
## if(run.simuls){
##     write_fst(simres,path=file.fst)
## } else {
##     simres <- read_fst(file.fst,as.data.table=TRUE)
## }
simres <- NMreadSim("simulate-results/NMsim_xgxr021_noname_paths.rds")
```

We plot population and individual predictions as the simulations of
(in this case) the typical subject and one simulated subject. The variable called `Y` is the individual prediction plus residual variability.
paper. The code is included to show that the results from `NMsim` are ready to be plotted. The main reason data is transformed to long format (`melt`) is to get ggplot2 to generate the legend automatically.

```{r,eval=TRUE}
datl <- as.data.table(simres) |>
    melt(measure.vars=cc(PRED,IPRED,Y))

ggplot(datl,aes(TIME,value,colour=variable))+
    geom_line(data=function(x)x[variable!="Y"])+
    geom_point(data=function(x)x[variable=="Y"])+
    labs(x="Hours since first dose",y="Concentration (ng/mL)")
```

The reason we can plot a simulation with residual variability is that the control stream includes a variable `Y` defined with residual variability in `$ERROR`:
```
  Y=F+F*ERR(1)+ERR(2)
```
More on residual variability in case you don't have such a line later
in this paper.

## What happened?
`NMsim` uses automation tools from `NMdata` to 

* Save the data in a Nonmem-friendly format 
* Create a simulation control stream based on the referenced input control stream
* Update the initial values based with estimated values 
* Modify input data-related sections for reading input siulation data 
* Modify output table file names and paths to generate simulation output tables
* Run Nonmem
* Read output tables and combine them with input data into one data object

The generated files have all been stored in a folder called `NMsim` next to the estimation control stream. More on that shortly in the section "A few basic additional arguments to `NMsim`".

Let's see the first few lines of the returned object:

```{r}
simres[1:3]
```

Notice a few things about the returned data:

* All columns from the output tables defined in the input control stream are there. We will soon learn how to modify this
* Input data columns are there too (like `trt`).
* Additional columns (`model` and `nmout`) may be familiar to `NMdata` users.
* We will soon learn where the "_noname" in the `model` column comes from.

## How to (re-)read the simulation results
`NMsim()` creates an `.rds` file with information about where all the results are stored. It is named based on the model name and the `name.sim` argument. NMsim provides the path to it in the console at every run. You can also specify exactly where it should be stored and the file name using the `file.res` argument. Run the function `NMreadSim` on this file to (re-)read all the simulation results. 

`NMreadSim` also supports the `wait` argument making it wait for all the simulation results to be available in case you submitted a large simulation to a cluster and want to continue your execution only when it's all done.

## Multiple models
Before we continue with that model, we want to compare a simulation based on this model to another model we are considering. `NMsim` can do this and collect the data into one object:

```{r,sim-twomodels,eval=FALSE}
files.2.mod <- file.project(c("nonmem/xgxr021.mod","nonmem/xgxr114.mod"))
simres.2models <- NMsim(file.mod=files.2.mod,
                        data=dat.sim
                        )
```

```{r,include=FALSE,eval=TRUE}
file.fst <- "simulate-results/simres_2models.fst"
if(run.simuls){
    write_fst(simres.2models,path=file.fst)
} else {
    simres.2models <- read_fst(file.fst,as.data.table=TRUE)
}
```
In case multiple models are provided, `NMsim` simply loops over them. It does collect all the results, and we can use the `model` column to separate the two simulations. Since we are so far just simulating on subject with each model, it makes litlle sense to compare individual preditions. We just plot the population prediction (`PRED`):

```{r,eval=TRUE}
ggplot(simres.2models,aes(TIME,PRED,colour=model))+geom_line()+
    facet_wrap(c("trt"),scales="free")
```

For simplicity, we shall show the rest of the examples for just one
model. Any of them could be run on multiple models the same way as
shown above.

<!-- subproblems vs repetition of input data -->

## A few basic additional arguments to `NMsim`
The first couple of examples were run with the bare minimum of
arguments - estimation control stream and simulation data set. You are
obviously encouraged to read the help of `NMsim` to learn about the
many useful features it has, and you will learn some more in this
vignette. But there are a few arguments that you should learn about at
this point already. Here they are:

* `dir.sims` Path to a folder in which all generated files will be
  stored. Use this to avoid `NMsim` will write into the directories
  where your estimation models are. They may not belong there at
  all. You may want to separate the model development step from the
  post-processing step. You are encouraged to explore what `NMsim`
  leaves in this directory (you will find fully reproducible
  simulation Nonmem runs including simulation input data).
* `dir.res` while `dir.sims` contains all the Nonmem files, you can
  specify a separate directory just for compressed results. The `rds`
  files containing information about where the simulations were
  performed will also be saved here. This means that as soon as
  results have been read once, all contents of `dir.sims` can be
  purged without loss of critical data. This can save a lot of disk
  space.
* `name.sim` Give your simulation a meaningfull name. We did not do
  this above, so `NMsim` called it "noname".
* `table.vars` Very important. This redefines the output table section
  from the estimation control stream to the simulation control
  stream. The estimation control stream may have too many variables
  printed (which will make Nonmem slow), or it may not have some that
  are useful for the simulation analysis. See how this is used
  below. Once you get used to this argument, you will use it very
  frequently.
* `wait` Wait for simulation to be done and return the resulting data?
  If not a path to the rds file to be read with `NMreadSim` will be
  returned.
* `reuse.results` If `TRUE` and results are found on file, those will
  be read instead of rerunning the simulation.
* `seed` A numeric value that will be used in Nonmem's $SIMULATION section


You will learn about a few more arguments in the next examples.



## More subjects and prediction intervals
To create a prediction interval based on the selected model, we need
to simulate multiple new subjects. There are two ways to easily obtain
that. One is to repeat (`rbind`) the simulation input dataset, one
repetetion per new subject, and then update the `ID` column to get
distinct subjects. 


### Multiple subjects created in simulation input data
The follwing shows how one could generate 1000 subjects using
`data.table`. (I use `data.table` a lot, if you can provide a good way
to do this without, I am happy to include that).

```{r,eval=FALSE,collapse=TRUE,class.source = "fold-hide"}
dat.sim.1000 <- NMdata::egdt(
                            as.data.table(dat.sim)[,!("ID")]
                           ,
                            data.table(ID=1:1000)
                        )
dat.sim.1000[,ID:=.GRP,by=.(ID,trt)]
## order with respect to new IDs
setorder(dat.sim.1000,trt,ID,TIME,EVID)
## check dataset
NMcheckData(dat.sim.1000,type.data="sim")
```

We now simulate 1000 subjects by plugging in this data object:
```{r,sim-1000-data,eval=FALSE}
simres.n1000.1 <- NMsim(file.mod=file.mod,
                        data=dat.sim.1000,
                        dir.sims="~/NMsim_vignette", ## where to store simulation files
                        name.sim="N1000_datarep"
                        )
```

```{r,include=FALSE,eval=TRUE}
## file.fst <- "simulate-results/simres_n1000_1.fst"
## if(run.simuls){
##     write_fst(simres.n1000.1,path=file.fst)
## } else {
##     simres.n1000.1 <- read_fst(file.fst,as.data.table=TRUE)
## }
simres.n1000.1 <- NMreadSim("simulate-results/NMsim_xgxr021_N1000_datarep_paths.rds")
```

### Multiple subjects generated by Nonmem
The other way to simulate multiple subjects is making use of Nonmem's `SUBPROBLEMS` simulation feature which makes Nonmem rerun the simulation the specified number of times. Notice that to do this, we use the `dat.sim` data without the 1000 replications. We then make use of the NMREP column generated by `NMdata::NMscanData` to redefine the `ID` column:


```{r,sim-1000-subproblem,eval=FALSE}
simres.n1000.2 <- NMsim(file.mod=file.mod,
                        data=dat.sim,
                        subproblems=1000
                        dir.sims="~/NMsim_vignette", ## where to store simulation files
                        dir.res="simulate-results", ## where to store simulation files
                        name.sim="N1000_subproblems",
                        )
simres.n1000.2 <- as.data.table(simres.n1000.2)[,ID:=.GRP,by=.(NMREP,ID,trt)]
```

```{r,include=FALSE,eval=TRUE}
file.fst <- "simulate-results/simres_n1000_2.fst"
if(run.simuls){
    write_fst(simres.n1000.2,path=file.fst)
} else {
    simres.n1000.2 <- read_fst(file.fst,as.data.table=TRUE)
}
```


The two approaches are computationally about equally fast, the most significant difference probably being in Nonmem reading a smaller or larger simulation input data file. Unless the input dataset becomes very large, it is merely a question of preference of the modeler which one to use. In a case where the simulated patients need different dosing or sample schedules, the manual construction of the data is needed - because it's not a straightforward replication.

### The prediction interval
We now plot a prediction interval - in this case based on the results of the simulation using `SUBPROBLEMS`; this makes no difference to how to derive the prediction interval.

```{r,eval=TRUE}
simres.pi <- as.data.table(simres.n1000.2)[,setNames(as.list(quantile(IPRED,probs=c(.05,.5,.95))),cc(ll,median,ul)),
                                           by=.(trt,TIME)]
simres.pi$type <- "pi"
simres.pi$pi.cover <- "90%"

p.pi.typ <- ggplot(simres.pi,aes(TIME,fill=trt))+
    geom_ribbon(aes(ymin=ll,ymax=ul,alpha=pi.cover))+
    geom_line(aes(y=median,colour=trt))+
    scale_alpha_manual(values=c("90%"=.5))+
    labs(x="Hours since first dose",y="Concentration (ng/mL)")

p.pi.typ
```

## Read previously generated simulations
There is no need to save simulation results because they are already saved by `NMsim`. Instead, use arguments `dir.sims`, `dir.res` and `name.sim` to make sure to get a meaningful structure for the generated files. Then read the results with `NMreadSim()`.

```{r,eval=FALSE}
library(NMdata)
simres.n1000.1 <- NMreadSim("simulate-results/simres_n1000_1_paths.rds")
```

In fact, that is also what `NMsim` does once Nonmem has run.


