---
title: "Simulate with modified parameter values"
output:
rmarkdown::html_vignette:
    toc: true
Suggests: markdown
VignetteBuilder: knitr
vignette: >
  %\VignetteIndexEntry{VaryParams}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
- \usepackage{ae}
---

```{r,include = FALSE}
library(tidyvpc)
library(NMdata)
library(fst)
##knitr::opts_chunk$set(dev = "cairo_pdf")
knitr::opts_chunk$set(
                      collapse = TRUE
                     ,comment = "#>"
                     ,fig.width=7
                     ,cache=FALSE
                  )

## NMdataConf(dir.psn="/opt/psn")

## this changes data.table syntax. I think we can do without.
## knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
run.simuls <- FALSE
```


## Vary parameter values
Sometimes we want to simulate with some modification to the estimated model. NMsim can make such user-specified modifications to the model before simulating through the `list.sections` argument. 

The SAD study was run with a fast solution formulation. We want to see how a slower absorption rate would affect the PK prediction for the multiple dose regimen. In the model estimate, `TVKA=2.17`. We now try with a four times slower absorption:

```{r,eval=FALSE}
simres <- NMsim(file.mod=file.mod
               ,data=dat.sim
               ,dir.sims="~/NMsim_vignette" ## where to store simulation files
               ,seed=12345
                )

simres.slowabs <- NMsim(file.mod=file.mod,
                        data=dat.sim
                       ,dir.sims="~/NMsim_vignette" ## where to store simulation files
                       ,name.sim="slower_abs"
                       ,seed=12345
                       ,list.sections=list(PK=function(x)c(x,"TVKA=TVKA/4","KA=KA/4"))
                        )

```
```{r,include=FALSE,eval=TRUE}
file.fst <- "simulate-results/simres_slowabs.fst"
if(run.simuls){
    write_fst(simres.slowabs,path=file.fst)
} else {
    simres.slowabs <- read_fst(file.fst,as.data.table=TRUE)
}
simres <- read_fst("simulate-results/simres.fst",as.data.table=TRUE)
```

```{r,eval=TRUE}
rbind(simres,simres.slowabs) |>
    ggplot(aes(TIME,PRED,colour=model))+
    geom_line()

```

We used `list.sections` to modify the `$PK` section. We used it to
append two lines. We could use it to modify any section in the model,
and we could essentially do any modification. However, appending to
`$PK` or `$PRED` is simple and often both robust and flexible enough.

That was a very spcific analysis of one specific `KA` value. It is often more convenient to control the numeric changes to the model using the simulation input data set rather than hard-coding numerical values into `list.sections`. The following tries a number of fold changes to `KA`.

```{r,include=TRUE,eval=FALSE}
NMdataConf(as.fun="data.table")
dat.sim.varka <- egdt(dat.sim,data.table(KASCALE=c(1,4,10)))
dat.sim.varka[,ID:=.GRP,by=.(KASCALE,ID)]
setorder(dat.sim.varka,ID,TIME,EVID)

simres.varka <- NMsim(file.mod=file.mod,
                      data=dat.sim.varka
                     ,dir.sims="~/NMsim_vignette" ## where to store simulation files
                     ,name.sim="varka"
                     ,seed=12345
                     ,list.sections=list(PK=function(x)c(x,"TVKA=TVKA/KASCALE","KA=KA/KASCALE"))
                      )

```
```{r,include=FALSE,eval=TRUE}
file.fst <- "simulate-results/simres_varka.fst"
if(run.simuls){
    write_fst(simres.varka,path=file.fst)
} else {
    simres.varka <- read_fst(file.fst,as.data.table=TRUE)
}
```

```{r,eval=TRUE}
ggplot(simres.varka[simres.varka$EVID==2,],aes(TIME,PRED,colour=factor(KASCALE)))+geom_line()+labs(colour="Fold absorption prolongation")
```



<!-- ## Using NMsim to validate reimplementations of models -->

