---
title: "Simulate Known Subjects Using Emperical Bayes Estimates (Etas/Phis)"
output:
rmarkdown::html_vignette:
    toc: true
    code_folding: show
Suggests: markdown
VignetteBuilder: knitr
vignette: >
  %\VignetteIndexEntry{Known}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
- \usepackage{ae}
---

```{r,include = FALSE}
##knitr::opts_chunk$set(dev = "cairo_pdf")
knitr::opts_chunk$set(
                      collapse = TRUE
                     ,comment = "#>"
                     ,fig.width=7
                     ,cache=FALSE
                  )

## this changes data.table syntax. I think we can do without.
## knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

```{r,setup,include=F}
## library(devtools)
## unloadNamespace("NMsim")
## unloadNamespace("NMdata")

## load_all("~/wdirs/NMdata")
## load_all()
library(NMsim)
library(data.table)
library(NMdata)
## library(dplyr)
library(tibble)
library(ggplot2)
library(patchwork)
library(tracee)
## library(tidyr)
library(fst)
library(knitr)
NMdataConf(path.nonmem="/opt/NONMEM/nm75/run/nmfe75")
## NMdataConf(path.nonmem="/opt/nonmem/nm751/run/nmfe75")
## NMdataConf(dir.psn=NULL)

theme_set(theme_bw())
this.script <- "NMsim-methods.Rmd"
writeOutput <- TRUE
file.project <- function(...)file.path(system.file("examples",package="NMsim"),...)
## file.project <- function(...)file.path("../inst/examples",...)
## file.project <- function(...)file.path("~/wdirs/NMsim/inst/examples",...)

found.files <- list.files(file.project("nonmem/NMsim"),pattern="noname\\.(lst|xml|ext|cov|cor|coi|phi|msf|msfi|msfo|tab)",full.names=TRUE)
unlink(found.files)

run.simuls <- FALSE
NMdataConf(as.fun="data.table")

dat.sim <- read.csv(system.file("examples/derived/dat_sim1.csv",package="NMsim"))
```


Built `r Sys.Date()` using NMsim `r packageVersion("NMsim")`.


## Objectives
This vignettes aims at enabling you to use `NMsim` for the following purposes

* Simualation of known subjects (estimated random effects), 


## Prerequisites
You should have configured `NMsim` with the path to the Nonmem
installation and maybe also PSN (optional). See
[`NMsim-config.html`](https://philipdelff.github.io/NMsim/articles/NMsim-config.html). Don't
worry - it is very easy. 

You should be familiar with basic `NMsim` arguments as described in
[`NMsim-basics.html`](https://philipdelff.github.io/NMsim/articles/NMsim-basics.html). In
that vignette you should have learned to use the default simulation
method. This vignette will be using the same model and simulation
input data to demonstrate how to use additional methods and features of `NMsim`.


## Simulation of known subjects 
We sometimes want to simulate the already observed subjects. This
means we want to reuse the estimated random effects (ETA's) given the
subject ID's. `NMsim` has a method for this called `NMsim_EBE`. The restriction is that
all subjects (values of `ID`) in the simulation input data must have
been used in the estimation input data.

Let's think about that one more time before we go on. For
`NMsim_EBE` to work, `ID` values in the simulation data set must be
identical to `ID` values used in the estimation. This is how
`NMsim_EBE` will find the EBE's (ETAs) that define the subject. For
models estimated with FOCEI, `NMsim_EBE` will take the ETAs from the
`.phi` file produced by Nonmem. For SAEM/IMP-based estimation this
file cannot be used, and the user must have stored the ETAs in output
table files for the method to work. It does not matter where in the
table files they are, they just all need to be there. You can write
all of them like it is done in the example model provided with `NMsim`
called `xgxr032.mod`.

```
$TABLE ETAS(1:LAST) NOAPPEND NOPRINT FILE=xgxr032_etas.txt
```

So a "subject" essentially means a set of ETAs. Covariates must still
be provided in the input data set and can be modified if
wanted. Notice, `NMsim_EBE` does nothing to restore covariates - you
must define covariates in input data as needed.

## Individual dosing history, sample times or both
Individual simulations are useful for several purposes. Examples

* Reuse individual dosing history and use a new common sample
  scheme. This could be for homogeneous evaluation of exposure metrics
  such as Cmax, AUC, or just to show concentration-time profiles at
  identical sampling times based on a model estimated on data with
  heterogeneos sampling. Currently, there is no example of this in
  this vignette but see the other examples, and it should be pretty
  clear how to do this.

* Reuse dosing history and use individual sampling scheme from a
  different data set, e.g. PD.

* Simulate a new common dosing and sampling scheme to simulate already
  observed subjects on a new regimen. This is sometimes done if for
  one reason or the other the model is not considered reliable for
  simulation of new subjects but the individual parameter estimates
  are trusted. 
  
* Reuse a simulated population. One may prefer to reuse the same
  simulated subjects in multiple simulations for reproducibility and to have all difference
  between say simulation results of different regimens be driven by
  differences in the regimen, and not in the populations. This
  use of `NMsim_EBE` is on the todo list tog get it's own vignette
  [here](https://philipdelff.github.io/NMsim/articles/NMsim-ReuseSimSubjects.html).

## Simulate known subjects on a new dosing regimen and new sample schedule
The following code takes an already created simulation data set with a
single ID, and merges all other columns than `ID` onto the observed
`ID`s. That gives the same simulation data for all subjects. 

First thing, we decide on a model (an input control stream of an estimated model) to use for the example:
```{r}
file.mod <- system.file("examples/nonmem/xgxr021.mod",package="NMsim")
```

Simulation with known ETA values is only possible through NMsim's internal Nonmem execution method. That means, we need to provide the path to the Nonmem executable.

```{r,eval=TRUE}
NMdataConf(path.nonmem="/opt/NONMEM/nm75/run/nmfe75")
```

For a simulation data set we are using the multiple dose simulation data set with a loading dose generated in the [Creation of Simulation Data Sets](https://philipdelff.github.io/NMsim/articles/NMsim-DataCreate.html).


```{r,eval=FALSE}
## read model results just to extract the observed ID's. Notice, if
## your model uses covariates, those should be retrieved too.
res.mod <- NMscanData(file.mod,quiet=TRUE)
ids <- data.frame(ID=unique(res.mod$ID))

## Repeat the simulation data set for each ID and order accordingly
dat.sim <- read.csv(system.file("examples/derived/dat_sim1.csv",package="NMsim"))
## we will merge with all the observed subjects in stead of the single
## ID in the original data set.
dat.sim$ID <- NULL
dat.sim.known <- merge(ids,
                       dat.sim
                       )
setorder(dat.sim.known,ID,TIME,EVID)
## check data
NMcheckData(dat.sim.known,type.data="sim")

simres.known <- NMsim(file.mod=file.mod,
                      data=dat.sim.known,
                      method.sim=NMsim_EBE,
                      table.vars="PRED IPRED CL V2 KA",
                      name.sim="known1"
                      )
```

```{r,include=FALSE,eval=TRUE}
file.fst <- "simulate-results/xgxr021_known1_paths.rds"
simres.known <- NMreadSim(file.fst)
```

And the simulation results are plotted for each subject.
```{r,eval=TRUE}
ggplot(subset(simres.known,EVID==2),aes(TIME,IPRED,colour=factor(ID)))+
    geom_line()+
    theme(legend.position="none")
```

<!-- Show a plot of the individual parameters in estimate and in sim -->

### Simulate individual dosing history at new individual sampling times for a PK/PD dataset
We want to plot some PD data angainst
PK. However, PD was sampled differnetly than PK, and we want to
evaluate the individual predictions of the PK model at the individual
PD samplng times - reusing the individual dosing history.

Reading some example PD data:
```{r,pddata,include=TRUE,eval=TRUE}
pd <- readRDS(system.file("examples/data/xgxr_pd.rds",package="NMsim"))
## some code below makes use of data.table features so we make it a data.table
setDT(pd)
```

For a PK model without time-varying covariates, suggested steps 
to generate the data for the simulation are:

* Take dose records from PK model estimation input data (`pkdos`). Just keep necessary columns like `ID`, `TIME`, `EVID`, `CMT`, `AMT`, `ADDL`, `II`, and any necessary covariates
* Take PD data observation records (`pdsamples`). Just keep `ID`, `TIME`, and set `EVID=2`.
* Add a unique row identifier to `pdsamples` (an integer row counter, like `ROW=1:nrow(pdsamples)`)
* Stack (`rbind` for data.tables or `bind_rows` in tidyverse) `pkdos` and `pdsamples` to one data set (`pdsim`)
* In `pdsim`, set `DV=NA`
* Sort `pdsim` at least by `ID`, `TIME` and `EVID`. There could be more depending on trial design

In case of time-varying covariates, you can keep all data records from the PK data (without `DV`), but change observation records to simulation records (`EVID=2` instead of `EVID=0`).

```{r,known-pkpd,eval=TRUE}
## Take dose records from PK model estimation input data
pkres <- NMscanData(file.mod,quiet=TRUE)
pkdos <- pkres[EVID==1,.(ID, TIME, EVID, CMT, AMT)]
## Take PD data observation records (`pdsamples`)
pdsamples <- pd[EVID==0,.(ID,TIME,LIDV)]
## Stack `pkdos` and `pdsamples` to one data set (`pdsim`)
pdsim <- rbind(pkdos,pdsamples,fill=TRUE)
## Nonmem needs a DV column to run
pdsim[,DV:=NA]
## only include subjects that were included in the PK model
pdsim <- pdsim[ID%in%pkres$ID]
setorder(pdsim,ID,TIME,EVID)
```

Then run `NMsim` like this. We are renaming the prediction columns using Nonmem `$TABLE` statement not to confuse PK predictions and PD predictions.

```{r,known-pkpd-run,eval=FALSE}
simres.pksim <- NMsim(file.mod,
                      data=pdsim,
                      name.sim="pkpd",
                      method.sim=NMsim_EBE,
                      table.vars="PKIPRED=IPRED PKPRED=PRED"
                      )
```
```{r,include=FALSE,eval=TRUE}
file.fst <- "simulate-results/xgxr021_pkpd_paths.rds"
simres.pksim <- NMreadSim(file.fst)
```



```{r,eval=TRUE}
ggplot(simres.pksim[!is.na(LIDV)&!is.na(PKIPRED)],aes(PKIPRED,LIDV))+
    geom_point()+
    labs(x="Individual PK prediction",y="Observed PD value")
``` 


