

## reuse estimation data or supply new simulation input data?


## Notice: These features are only available in the NMsim development version 
In NMsim 0.0.4 and earlier versions NMsim required a data set to be
supplied using the `data` argument. The steps shown in this vignette
do work with those versions if you do that, but it is better to avoid
it. The way the new versions of NMsim does this if no data set is
supplied is safer because it does not process the input data at
all. It reruns the estimated model on the estimation input data set
the exactly as specified in the estimation control stream. Once one
reads the data set, applies filters like specified in Nonmem, renames
columns like Nonmem does, mistakes can be made. The new versions avoid
this risk.

## Including simulated residual error 
The simulations in this vignette are based on using a Nonmem variable `Y` 
evaluates prediction plus simulated residual error. The variable
does not need to be called `Y`, Whether the variability (variance or std. deviation) is estimated using `$SIGMA` or `$THETA` does not matter. One example of a combined error model (here, `$SIGMA` is used) is:

```
Y=F+F*ERR(1)+ERR(2)
```

## Running NMsim
Normally, the two main arguments to NMsim are the path to the input
control stream (`file.mod`) and the simulation input data set
(`data`). But if we leave out the the `data` argument, NMsim will
re-use the estimation data for the simulation. That is the simulation
we need for a VPC. We will use an example model included with NMsim:

```{r}
file.project <- function(...)file.path(system.file("examples",package="NMsim"),...)
file.mod <- file.project("nonmem/xgxr022.mod")

set.seed(43)
## notice the data argument is not used.
simres.vpc <- NMsim(file.mod,
                    ## we don't need much for the output table. Notice
                    ## Y is sim with residual variability
                    table.vars=c("PRED","IPRED", "Y"),
                    dir.sims="testOutput",
                    name.sim="vpc_01"
                   ,nsims=3
                   ,method.execute="nmsim"
                   ,path.nonmem=path.nonmem
                    )

```


```{r}
simdir <- file.path("simulations",paste(fnExtension(basename(file.mod),""),"vpc3",sep="_"))
simres <- NMscanMultiple(dir=simdir,quiet=TRUE,use.input=FALSE)
simres <- NMreadSim("simulations/xgxr0 )

cols.trans <- cc(Y,IPRED,PRED)
simres[,(cols.trans):=lapply(.SD,fun.trans),.SDcols=cols.trans]
dims(simres)

```




## Parallelized simulation
Running 500 or 1000 simulations may take a while if your data set is
large. The following example runs only ten repetitions in a Nonmem
runs, but it runs 100 of those. 


```{r}
set.seed(9857)    
simpaths <- NMsim(file.mod,
                  ## we don't need much for the output table. Notice
                  ## Y is sim with residual variability
                  table.vars="TIME PRED IPRED Y",
                  dir.sims="simulations",
                  name.sim="vpc3",
                  ## We do 1000 reps but its slow. So just 10 on
                  ## each run and 100 runs.
                  subproblems=10,
                  nsims=100
                  ## run on the cluster
                 ,sge=TRUE
                  )
```
